### 2013-12-12 inst/doc/manuscript.Rmd
a little writing on manuscript

* Why NeXML -- because Nexus standard isn't a standard.  Developers extend it arbitrarily to meet their needs, making it incompatible.  Cannot represent horizontal gene transfer, ambiguous use of symbols, cannot be extended reliably, no validator.  NeXML solves all of these problems [Vos et al] with a state-of-the-art data management format.

@@

### Reading NeXML

@@

```{r}
f <- system.file("examples", "trees.xml", package="RNeXML")
nex <- read.nexml("trees.xml")
```

The resulting `nex` object is an R object representation of the NeXML.

```{r}
phy <- get_trees(nex)
plot(phy)
```

Where we have used the standard plotting function from the `ape` package to plot the resulting `ape::phylo` object returned by `get_trees`.

@@

>>>>

* Why NeXML -- because Nexus standard isn't a standard.  Developers extend it arbitrarily to meet their needs, making it incompatible.  Cannot represent horizontal gene transfer, ambiguous use of symbols, cannot be extended reliably, no validator.  NeXML solves all of these problems [Vos _et al._ 2012] with a state-of-the-art data management format.

@@

## Working with phylogenetic trees

#### Read methods

@@

```{r}
f <- system.file("examples", "trees.xml", package="RNeXML")
nex <- nexml_read("trees.xml")
```

(For convenience `nexml_read` is also aliased as `read.nexml`, for users familiar with the somewhat analogous `read.nexus` function).   The resulting `nex` object is an R object representation of the NeXML.  From this object we can extract any phylogenies it contains in the `ape::phylo` format:

```{r}
phy <- get_trees(nex)
```
We can then leverage the rich suite of methods available from the R phylogenetics community (including coercion into other representations, such as `ouch` [@ouch] and `phylobase` [@phylobase] phylogeny formats). For instance, here we use the standard plotting function from the `ape` package to plot the resulting `ape::phylo` object returned by `get_trees`.

[@ouch]: http://cran.r-project.org/web/packages/ouch/
[@phylobase]: http://cran.r-project.org/web/packages/phylobase

``{r}
plot(phy)
```

@@

<!-- have we already said all this above?-->
The ability to read in phylogenetic trees in the `nexml` format opens up a wide and rapidly growing array of resources to the R user.  For instance, a user can access all the phylogenies available in TreeBASE through the `nexml` format.  `RNeXML` can read directly from a URL:

```{r eval=FALSE}
nex = nexml_read("https://raw.github.com/rvosa/supertreebase/master/data/treebase/S100.xml")
```

Previously, this was possible using the `treebase` package [@treebase] for R, which relied on the nexus parser provided in the `ape` package.  Due to inconsistencies present in the Nexus standard [Vos _et al._ 2012], a significant number <!-- State explicitly --> of TreeBASE nexus files could not be read in this way.  Using the more precisely defined NEXML standard, every valid NeXML file on TreeBASE can now successfully be parsed.

#### Write methods
